Model validation plots

### 2021-08-18 doc/validate_models.Rmd
add supplemental to choose correct knot number #92

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
## target knits Rmds in their own session, so load libraries here.
library(targets)
library(statmod)
library(tidyverse)
library(gratia)
library(mgcv)
library(dlnm)
library(arm)
```

@@

- increase knots for some models (https://github.com/BrunaLab/HeliconiaDemography/issues/92)
- check diagnostics for individual-level random effects.  Doesn't look good, but not sure how to interpret or what to do.

@@

```{r echo=TRUE, eval=TRUE}
s_1ha_res <- qresid(s_1ha)
par(mfrow = c(2,2))
#histogram
plot(density(s_1ha_res))
#QQ plot
qqnorm(s_1ha_res); qqline(s_1ha_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(s_1ha), residuals(s_1ha, type = "response"))
```

@@

```{r eval=TRUE}
# looking for near zero edf
gam(s_1ha_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(s_1ha)) %>%
  edf() #fine

gam(s_1ha_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(s_1ha)) %>%
  edf()#fine
```

@@

```{r echo=TRUE}
s_cf_res <- qresid(s_cf)
par(mfrow = c(2,2))
#histogram
plot(density(s_cf_res))
#QQ plot
qqnorm(s_cf_res); qqline(s_cf_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(s_cf), residuals(s_cf, type = "response"))
```
### Basis dimensions

@@

```{r}
# looking for near zero edf
gam(s_cf_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(s_cf)) #fine

gam(s_cf_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(s_cf)) 
```

@@

```{r}
g_1ha_res <- residuals(g_1ha)
# looking for near zero edf
gam(g_1ha_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(g_1ha)) #fine

gam(g_1ha_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(g_1ha)) #fine
```

@@

```{r}
g_cf_res <- residuals(g_cf)
# looking for near zero edf
gam(g_cf_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(g_cf)) #fine

gam(g_cf_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(g_cf)) #fine
```

@@

```{r}
f_1ha_res <- qresid(f_1ha)
par(mfrow = c(2,2))
#histogram
plot(density(f_1ha_res))
#QQ plot
qqnorm(f_1ha_res); qqline(f_1ha_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(f_1ha), residuals(f_1ha, type = "response"))
```

@@

```{r}
# looking for near zero edf
gam(f_1ha_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(f_1ha)) #fine

gam(f_1ha_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(f_1ha)) #fine
```

@@

```{r}
f_cf_res <- qresid(f_cf)
par(mfrow = c(2,2))
#histogram
plot(density(f_cf_res))
#QQ plot
qqnorm(f_cf_res); qqline(f_cf_res)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(f_cf), residuals(f_cf, type = "response"))
```

@@

```{r}
# looking for near zero edf
gam(f_cf_res ~ s(log_size_prev, k = 10, bs = "cs"), gamma = 1.4,
    data = model.frame(f_cf)) #fine

gam(f_cf_res ~ s(spei_history, L,
      bs = "cb",
      k = c(3, 35),
      xt = list(bs = "cr")),
    gamma = 1.4,
    select = TRUE,
    data = model.frame(f_cf)) #fine
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
## target knits Rmds in their own session, so load libraries here.
library(targets)
library(statmod)
library(tidyverse)
library(gratia)
library(mgcv)
library(dlnm)
library(arm)
source(here::here("R", "utils.R"))
```

@@

- check that # of knots are adequate
- check diagnostics for individual-level random effects.  Doesn't look good, but not sure how to interpret or what to do.

@@

```{r echo=TRUE, eval=TRUE}
s_1ha_qres <- qresid(s_1ha)
par(mfrow = c(2,2))
#histogram
plot(density(s_1ha_qres))
#QQ plot
qqnorm(s_1ha_qres); qqline(s_1ha_qres)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(s_1ha), residuals(s_1ha, type = "response"))
```

@@

```{r eval=TRUE}
check_res_edf(s_1ha)
```

The crossbasis smooth possibly needs more knots.

```{r}
# plot_lag_slice(s_1ha, "spei_history", lag =  0) + labs(title = "surv, 1ha")
```

Here you can see how the CIs hardly overlap 0, but it could be an artifact of the line not being allowed to be wiggly enough.

@@

```{r echo=TRUE}
s_cf_qres <- qresid(s_cf)
par(mfrow = c(2,2))
#histogram
plot(density(s_cf_qres))
#QQ plot
qqnorm(s_cf_qres); qqline(s_cf_qres)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(s_cf), residuals(s_cf, type = "response"))
```

@@

```{r}
# looking for near zero edf
check_res_edf(s_cf)
```

@@

```{r}
check_res_edf(g_1ha)
```

@@

```{r}
check_res_edf(g_cf)
```

```{r}
# plot_lag_slice(g_1ha, "spei_history", lag = 33)
```

hmm... CIs are **super** narrow here, but effect size is **tiny**

@@

```{r}
f_1ha_qres <- qresid(f_1ha)
par(mfrow = c(2,2))
#histogram
plot(density(f_1ha_qres))
#QQ plot
qqnorm(f_1ha_qres); qqline(f_1ha_qres)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(f_1ha), residuals(f_1ha, type = "response"))
```

@@

```{r}
# looking for near zero edf
check_res_edf(f_1ha)
```

@@

```{r}
f_cf_qres <- qresid(f_cf)
par(mfrow = c(2,2))
#histogram
plot(density(f_cf_qres))
#QQ plot
qqnorm(f_cf_qres); qqline(f_cf_qres)
#Fitted vs. residuals--binned residuals plot
binnedplot(fitted(f_cf), residuals(f_cf, type = "response"))
```

@@

```{r}
# looking for near zero edf
check_res_edf(f_cf)
```
